---
title: "6101 Project 1"
author: Jeffrey Hu, Aaron Rock, Sanchit Vijay
date: 9/27/22 to 10/27/22
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# Make sure that install.packages("matrixStats"), install.packages("reshape2"), install.packages("reshape") are installed

library(ezids)
library(ggplot2)
library(tidyverse)
library(MASS)
library(reshape)
library(dplyr)
library(matrixStats)
library(zoo)
library(ggfortify)
library(reshape2)
library(hexbin)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Import the data
```{r}
#First, load the file
zillow = data.frame(read.csv("price.csv"))
head(zillow)
```

## Remove blank pricing data
``` {r}
# Remove empty rows, NA's are only present if no data was collected for the location
zillow = na.omit(zillow)
```

## Subset the Zillow data into east coast and west coast and annual subsets of the regions.
``` {r}
# going through and discarding states that are not defined by the Uited States as West or East coast states

eastcoast = subset(zillow, State == 'CT' | State == 'NY' | State == 'VA' | State == 'FL' | State == 'NJ' | State == 'NC' |State == 'SC' | State == 'GA' | State == 'DE' | State == 'MA' | State == 'MD' | State == 'ME' | State == 'RI' | State == 'NH')
eastcoast_annual = eastcoast[,c(1:81)]
eastcoast_annual = eastcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
eastcoast_annual = eastcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
eastcoast_annual = eastcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
eastcoast_annual = eastcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
eastcoast_annual = eastcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
eastcoast_annual = eastcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
eastcoast_annual = eastcoast_annual[-c(14,15)]

westcoast = subset(zillow, State == 'CA' | State == 'OR' | State == 'WA' | State == 'HI' | State == 'AK')

westcoast_annual = westcoast[,c(1:81)]
westcoast_annual = westcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
westcoast_annual = westcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
westcoast_annual = westcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
westcoast_annual = westcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
westcoast_annual = westcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
westcoast_annual = westcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
westcoast_annual = westcoast_annual[-c(14,15)]

westcoast = westcoast[-c(2,3,4,5,6)]

westAnn = westcoast_annual
westAnn = westAnn[-c(2,3,4,5,6)]

eastAnn = eastcoast_annual
eastAnn = eastAnn[-c(2,3,4,5,6)]
```

## Find the difference between of rent price and add it as a column AND add average price of rent over the years column
```{r}
# use dplyr package to create new columns Price.Diff and Avg.Price
eastcoast_annual <- eastcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

westcoast_annual <- westcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

eastcoast_annual <- eastcoast_annual %>%
     mutate(Avg.Price = rowMeans(eastcoast_annual[7:13]))

westcoast_annual <- westcoast_annual %>%
     mutate(Avg.Price = rowMeans(westcoast_annual[7:13]))
```




## Clean Data of Outliers
```{r,results='asis'}
# clean the data
westcoast_annualClean = outlierKD2(westcoast_annual, November.2010, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, November.2016, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Avg.Price, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Price.Diff, rm = TRUE)

eastcoast_annualClean = outlierKD2(eastcoast_annual, November.2010, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, November.2016, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Avg.Price, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Price.Diff, rm = TRUE)
```

## Transpose Data Correctly
```{r}
rownames(westAnn) = westAnn$City.Code
westAnn = westAnn[c(-1)]
westAnn = as.data.frame(t(westAnn))
names(westAnn)<-str_replace_all(names(westAnn), c(" " = "" , "," = ""))
westAnn <- cbind(Date = rownames(westAnn), westAnn)
rownames(westAnn) <- 1:nrow(westAnn)
head(westAnn)


rownames(eastAnn) = eastAnn$City.Code
eastAnn = eastAnn[c(-1)]
eastAnn = as.data.frame(t(eastAnn))
names(eastAnn)<-str_replace_all(names(eastAnn), c(" " = "" , "," = ""))
eastAnn <- cbind(Date = rownames(eastAnn), eastAnn)
rownames(eastAnn) <- 1:nrow(eastAnn)
head(eastAnn)
```



## Summary of the two regions annually
``` {r, echo=T}
# getting the summaries to view basic stats of the data
summary(eastcoast_annual[7:15])
summary(westcoast_annual[7:15])
xkablesummary(westcoast_annual[7:15], title="West Coast Summary Annually")
xkablesummary(eastcoast_annual[7:15], title="East Coast Summary Annually")

xkablesummary(westcoast_annual[c(7,10,13:15)], title="West Coast Summary Annually")
xkablesummary(eastcoast_annual[c(7,10,13:15)], title="East Coast Summary Annually")
```

## Finding the Outliers to Help with Graphs
```{r}
# checking outliers
outWest2010 <- boxplot.stats(westcoast_annual$November.2010)$out
outWest2016 <- boxplot.stats(westcoast_annual$November.2016)$out
outEast2010 <- boxplot.stats(eastcoast_annual$November.2010)$out
outEast2016 <- boxplot.stats(eastcoast_annual$November.2016)$out
min(outWest2010)
min(outWest2016)
min(outEast2010)
min(outEast2016)
```

## West Coast Annual Novemeber 2010 Start and November 2016 End Histogram
```{r}
library(ggplot2)
ggplot(westcoast_annualClean, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2010", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2010", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2016", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annualClean, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2016", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 70) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 70) +
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")

colors <- c("November.2010" = "red", "November.2016" = "blue")

ggplot(westcoast_annualClean) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)
```

## East Coast Annual Novemeber 2010 Start and November 2016 End Histogram
```{r}
ggplot(eastcoast_annual, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annualClean, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annual, aes(November.2016)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annualClean, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2016", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annual) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 40) +
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")

colors <- c("November.2010" = "red", "November.2016" = "blue")

ggplot(eastcoast_annualClean) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)
```

## Boxplot of West Coast Zillow Rental Prices 2010-2016
```{r}
ggplot(westcoast_annual, aes(State, November.2010, color=State)) + geom_boxplot() + labs(title = "West Coast States November 2010 Rent Prices with Outliers")

ggplot(westcoast_annualClean, aes(State, November.2010, color=State)) + geom_boxplot()+labs(title = "West Coast States November 2010 Rent Prices without Outliers")

ggplot(westcoast_annualClean, aes(x=November.2010)) + geom_boxplot(color="blue", outlier.shape = 9, outlier.size = 2, outlier.color = "blue") + labs(title="Entire West Coast November 2010 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))

ggplot(westcoast_annual, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "West Coast States November 2016 Rent Prices with Outliers")

ggplot(westcoast_annualClean, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "West Coast States November 2016 Rent Prices without Outliers")

ggplot(westcoast_annualClean, aes(x=November.2016)) + geom_boxplot(color="blue", outlier.shape = 9, outlier.size = 2, outlier.color = "blue") + labs(title="Entire West Coast November 2016 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))
```

## Boxplot of East Coast Zillow Rental Prices 2010-2016
```{r}
ggplot(eastcoast_annual, aes(State, November.2010, color=State)) + geom_boxplot() + labs(title = "East Coast States November 2010 Rent Prices with Outliers")

ggplot(eastcoast_annualClean, aes(State, November.2010, color=State)) + geom_boxplot()+labs(title = "East Coast States November 2010 Rent Prices without Outliers")

ggplot(eastcoast_annualClean, aes(x=November.2010)) + geom_boxplot(color="red", outlier.shape = 9, outlier.size = 2, outlier.color = "red") + labs(title="Entire East Coast November 2010 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))

ggplot(eastcoast_annual, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "East Coast States November 2016 Rent Prices with Outliers")

ggplot(eastcoast_annualClean, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "East Coast States November 2016 Rent Prices without Outliers")

ggplot(eastcoast_annualClean, aes(x=November.2016)) + geom_boxplot(color="red", outlier.shape = 9, outlier.size = 2, outlier.color = "red") + labs(title="Entire East Coast November 2016 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))
```

## Scatter plot of West Coast
```{r, results='asis'}
ggplot(westcoast_annual, aes(November.2010,November.2016, color=State)) + geom_point() + labs(title="West Coast November 2010 and November 2016 with Outliers", x = "November.2010 Rent Prices", y = "November.2016 Rent Prices")
ggplot(westcoast_annualClean, aes(November.2010,November.2016, color=State)) + geom_point() + labs(title="West Coast November 2010 and November 2016 without Outliers", x = "November.2010 Rent Prices", y = "November.2016 Rent Prices")
```

## Scatter plot of East Coast
```{r, results='asis'}
ggplot(eastcoast_annual, aes(November.2010,November.2016, color=State)) + geom_point() + labs(title="East Coast November 2010 and November 2016 with Outliers", x = "November.2010 Rent Prices", y = "November.2016 Rent Prices")
ggplot(eastcoast_annualClean, aes(November.2010,November.2016, color=State)) + geom_point() + labs(title="East Coast November 2010 and November 2016 without Outliers", x = "November.2010 Rent Prices", y = "November.2016 Rent Prices")
```

## QQ Plots without Outliers
```{r, results='asis'}
qqnorm(westcoast_annualClean$November.2010, main="QQ-Plot West Coast November 2010 without Outliers")
qqline(westcoast_annualClean$November.2010)

qqnorm(westcoast_annualClean$November.2016, main="QQ-Plot West Coast November 2016 without Outliers")
qqline(westcoast_annualClean$November.2016)

qqnorm(eastcoast_annualClean$November.2010, main="QQ-Plot East Coast November 2010 without Outliers")
qqline(eastcoast_annualClean$November.2010)

qqnorm(eastcoast_annualClean$November.2016, main="QQ-Plot East Coast November 2016 without Outliers")
qqline(eastcoast_annualClean$November.2016)
```
## QQ Plots with Outliers
```{r, results='asis'}
qqnorm(westcoast_annual$November.2010, main="QQ-Plot West Coast November 2010 with Outliers")
qqline(westcoast_annual$November.2010)

qqnorm(westcoast_annual$November.2016, main="QQ-Plot West Coast November 2016 with Outliers")
qqline(westcoast_annual$November.2016)

qqnorm(eastcoast_annual$November.2010, main="QQ-Plot East Coast November 2010 with Outliers")
qqline(eastcoast_annual$November.2010)

qqnorm(eastcoast_annual$November.2016, main="QQ-Plot East Coast November 2016 with Outliers")
qqline(eastcoast_annual$November.2016)
```



## Format Date
```{r, results='asis'}
westAnn$Date = gsub('\\.',' ',westAnn$Date)
westAnn <- westAnn %>%
     mutate(Avg.Price = rowMeans(westAnn[2:1222]))
westAnn$Max = rowMaxs(as.matrix(westAnn[2:1222]))
westAnn$Min = rowMins(as.matrix(westAnn[2:1222]))
ggplot(westAnn, aes(Date, Avg.Price)) + geom_point() + labs(title="Average Price of the Years on the West Coast")


eastAnn$Date = gsub('\\.',' ',eastAnn$Date)
eastAnn <- eastAnn %>%
     mutate(Avg.Price = rowMeans(eastAnn[2:3989]))
eastAnn$Max = rowMaxs(as.matrix(eastAnn[2:3989]))
eastAnn$Min = rowMins(as.matrix(eastAnn[2:3989]))
ggplot(eastAnn, aes(Date, Avg.Price)) + geom_point() + labs(title="Average Price of the Years on the East Coast")



westAnnAvg = westAnn[-c(2:1222)]
eastAnnAvg = eastAnn[-c(2:3989)]
```

## T Test of First and Last Month of East and  West Coast
```{r,results='asis'}
loadPkg("BSDA") # for z.test
west2010_80 = t.test(x=westcoast_annualClean$November.2010, conf.level = 0.80)
west2010_80
west2016_80 = t.test(x=westcoast_annualClean$November.2016, conf.level = 0.80)
west2016_80

west2010_99 = t.test(x=westcoast_annualClean$November.2010, conf.level = 0.99)
west2010_99
west2016_99 = t.test(x=westcoast_annualClean$November.2016, conf.level = 0.99)
west2016_99

east2010_80 = t.test(x=eastcoast_annualClean$November.2010, conf.level = 0.80)
east2010_80
east2016_80 = t.test(x=eastcoast_annualClean$November.2016, conf.level = 0.80)
east2016_80

east2010_99 = t.test(x=eastcoast_annualClean$November.2010, conf.level = 0.99)
east2010_99
east2016_99 = t.test(x=eastcoast_annualClean$November.2016, conf.level = 0.99)
east2016_99
```

\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
Start of Cities of Interest File
\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\


## Load the file
```{r}
zillow_prices = data.frame(read.csv("price.csv"))
zillow_focus = zillow_prices[c(1:5,7,13,17,23,24,26,113,162),]
zillow_focus = zillow_focus[c(-1,-3,-4,-5,-6)]
zillow_focus
```


## Clean and Transpose
Our cities of interest are NYC (Queens) (only data from December 2011 Onwards), LA, PA, Houston (Harris), Chicago (Cook), Dallas, Las Vegas, San Fran, Detroit (Wayne), Roanoke, Richmond, District of Columbia, Baltimore, Seattle(King). Here, we clean data set and transpose the dates/cities
```{r, results="markup"}
rownames(zillow_focus) = zillow_focus$City
zillow_focus = zillow_focus[c(-1)]
zillow_focus = as.data.frame(t(zillow_focus))
names(zillow_focus)<-str_replace_all(names(zillow_focus), c(" " = "" , "," = ""))
head(zillow_focus)
str(zillow_focus)
```

Turn date into a column
```{r, results="markup"}
zillow_focus <- cbind(Date = rownames(zillow_focus), zillow_focus)
rownames(zillow_focus) <- 1:nrow(zillow_focus)
head(zillow_focus)
```


## Convert the Date column from chr to date, and Total
Here, we are converting strings to dates and to %d/%m/%Y form with 01 for %d
```{r, results="markup"}
zillow_focus_test = zillow_focus
zillow_focus_test$Date <- str_replace_all(zillow_focus_test$Date, "[.]", " ")
zillow_focus_test$Month <- str_extract(zillow_focus_test$Date, "(\\w+)")
zillow_focus_test$Year <- str_extract(zillow_focus_test$Date, "\\w+$")
tmp_date = paste(match(zillow_focus_test$Month, month.name), zillow_focus_test$Year, sep="/")
tmp_date = paste("01", tmp_date, sep="/")
tmp_date = as.Date(tmp_date, "%d/%m/%Y")
zillow_focus$Date <- tmp_date
head(zillow_focus)
str(zillow_focus)
```


Add a total column, not including New York (which has NAs)
```{r, results="markup"}
zillow_total = zillow_focus
zillow_total$Total <- rowSums(zillow_total[3:14])
head(zillow_total)
```


## Line plot of total
Line plots of time series, one with only time
```{r}
ggplot(zillow_total, aes(x=Date, y=Total)) + geom_point() + geom_line() + labs(x="Time", y="Total Rent Prices", title="Plot of City Rent Change over Time")
```


## Plots of the Cities and Prices
Melt the ggplot by the cities
```{r, results="markup"}
zillow_melt <- melt(zillow_focus, id = "Date")
names(zillow_melt)[2] <- "City"
names(zillow_melt)[3] <- "Prices"
head(zillow_melt)
```


I'm not sure if histograms would be meaningful in the context of my part, but here's one of all the cities melted together, ignoring New York again. We can do more if you guys believe it's necessary. There is a huge skew because the bigger/more expensive cities chosen would have higher rent prices due to a difference in the cost of living.
I'll also do the Mean/Median/Quartiles of the overall plots here.
```{r, results="markup"}
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_histogram(binwidth=300)
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_boxplot()
summary(zillow_melt$Prices)
```

Lines of the time series over their changes.
```{r}
ggplot(zillow_melt, aes(x=Date, y=Prices, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Rent Price", title="Plot of City Rent Prices over Time")
```


## First Order change plots
Find the first order change, difference between each month
```{r, results="markup"}
zillow_change <- as.data.frame(lapply(zillow_focus[2:14], diff, lag=1))
zillow_change$Date <- zillow_focus$Date[2:75]
zillow_change <- zillow_change %>% dplyr::select(Date, everything())
head(zillow_change)
```


Melt, and line graph of the changes for each city.
The graph is kind of hard to visualize, but it looks like San Francisco had the highest change from time to time.
```{r}
zillow_cmelt <- melt(zillow_change, id = "Date")
names(zillow_cmelt)[2] <- "City"
names(zillow_cmelt)[3] <- "PriceChange"
ggplot(zillow_cmelt, aes(x=Date, y=PriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Price Change", title="Plot of City Rent Change over Time")
```

Cumulative change, to make the trends more clear in respect to each other
```{r, results="markup"}
zillow_cumulative <- as.data.frame(cumsum(zillow_change[2:14]))
zillow_cumulative$Date <- zillow_focus$Date[2:75]
zillow_cumulative <- zillow_cumulative %>% dplyr::select(Date, everything())
head(zillow_cumulative)
```


Another graph, this time of cumulative change. San Francisco had the greatest change, while Seattle had the second largest. Detroit had the most negative change, falling below 0 overall.
```{r}
zillow_cumumelt <- melt(zillow_cumulative, id = "Date")
names(zillow_cumumelt)[2] <- "City"
names(zillow_cumumelt)[3] <- "CumulativePriceChange"
ggplot(zillow_cumumelt, aes(x=Date, y=CumulativePriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumulative Change", title="Plot of Cumulative City Rent Change over Time")
```


## First Order change, as a percentage proportion of the total rent
Find the first order change as a Percent
```{r, results="markup"}
zillow_proportion <- zillow_change[2:14]/zillow_focus[c(2:75), 2:14]*100
zillow_proportion$Date <- zillow_focus$Date[2:75]
zillow_proportion <- zillow_proportion %>% dplyr::select(Date, everything())
head(zillow_proportion)
```

It looks like Richmond has a pretty varied proportional change. It's still hard to make anything out of it, so let's move to the cumulative proportional change again.
```{r}
zillow_pmelt <- melt(zillow_proportion, id = "Date")
names(zillow_pmelt)[2] <- "City"
names(zillow_pmelt)[3] <- "ProportionChange"
ggplot(zillow_pmelt, aes(x=Date, y=ProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Proportional Change", title="Plot of City Rent Change over Time")
```


Quick histogram of the distribution of proportional changes. The proportional changes over time look pretty normal taken together as a whole, with maybe a slight right skew, which makes sense as we would guess that rent prices should generally increase over time with inflation and what not. Remember, these changes are in percentages. There are outliers on both sides of the boxplot, which matches what we see in the histogram. Here are the measures of variance and qqnorm plots as well
```{r, results="markup"}
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_histogram() + labs(x = "Proportional Change in Percentages")
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_boxplot()
qqnorm(zillow_pmelt$ProportionChange)
summary(zillow_pmelt$ProportionChange)
var(zillow_pmelt$ProportionChange, na.rm=TRUE)
sd(zillow_pmelt$ProportionChange, na.rm=TRUE)
```

Just making the cumulative melt for proportions now. We will drop New York before plotting the values, as there are too many NAs.

```{r, results="markup"}
zillow_pdrop <- zillow_proportion[3:14]
zillow_pcumulative <- as.data.frame(cumsum(zillow_pdrop))
zillow_pcumulative$Date <- zillow_focus$Date[2:75]
zillow_pcumulative <- zillow_pcumulative %>% dplyr::select(Date, everything())
head(zillow_pcumulative)
```

We will drop New York before plotting the values, as there are too many NAs.
Now, it looks like the proportional change of Seattle is a bit higher than San Francisco, and Richmond is up there too. Detroit still has the most negative proportional change overall.

```{r}
zillow_pcumumelt <- melt(zillow_pcumulative, id = "Date")
names(zillow_pcumumelt)[2] <- "City"
names(zillow_pcumumelt)[3] <- "CumulativeProportionChange"
ggplot(zillow_pcumumelt, aes(x=Date, y=CumulativeProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumu. Prop. Change", title="Plot of Cumulative City Rent Change over Time")
```

Box plot of each city's proportional change. As expected, San Francisco and Roanoke seem to have the largest ranges and variances. We are also dropping New York again.
```{r, fig.align="center", fig.width = 10, fig.height=5}
zillow_prodrop = zillow_proportion[-2]
zillow_pdmelt <- melt(zillow_prodrop, id = "Date")
names(zillow_pdmelt)[2] <- "City"
names(zillow_pdmelt)[3] <- "ProportionChange"
ggplot(zillow_pdmelt, aes(x=City, y=ProportionChange, color=City, na.rm=TRUE)) + geom_boxplot(outlier.shape=2, outlier.size=1)
```


## Anova Test and Post-hoc Tukey HSD
We will now anova test the proportional changes against each other, and see if the average change between years is different. Since we have a p-value less than our alpha, there are significant differences in changes, and thus, we look at the Post-hoc Tukey HSD. It looks like the pairs that have significantly different average proportional changes are [Detroit-LosAngeles], [Roanoke-LosAngeles], [SanFrancisco-Chicago], [Seattle-Chicago], [Richmond-Chicago], [Detroit-Houston], [Seattle-Philadelphia], [SanFrancisco-LasVegas], [Seattle-LasVegas], [Richmond-LasVegas], [Detroit-SanFrancisco], [Roanoke-SanFrancisco], [Seattle-Detroit], [Washington-Detroit], [Richmond-Detroit], [Baltimore-Seattle], [Roanoke-Seattle], and finally, [Roanoke-Richmond].
```{r, results="markup"}
pm_anova = aov(zillow_pdmelt$ProportionChange ~ zillow_pdmelt$City)
summary(pm_anova)
tukeyAoV <- TukeyHSD(pm_anova)
tukeyAoV
```

\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
Start of Total Population File
\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\

## State wise countplot  
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
ggplot(zillow, aes(x=State)) + geom_bar(colour="blue", fill="purple", alpha=0.6)
```

## Time-series plot for top 10 cities  
```{r, results='markup', fig.align="center", fig.width = 12, fig.height=5}
num_city = 10
values=head(zillow, num_city, )
values=data.frame(t(as.matrix(values[,7:81])))
colnames(values)=zillow[1:num_city,2]

date = seq(as.Date("2010/11/01"), as.Date("2017/01/31"), "month")
date = as.yearmon(date)
ts=zoo(values,order.by = date)
values=fortify(ts)
values$Index=as.Date(values$Index)

autoplot(ts,facets = NULL)+ geom_point(size=0.5) +
  theme_minimal()+
  labs(x="Time",y="Price")

```

## Pricing Distribution for California (Box plots)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
price = gather(data = zillow, "Month", "Price", 7:81, factor_key = T)

ggplot(price[price$State == "CA" & !is.na(price$Metro),]) +
  geom_boxplot(aes(x = fct_rev(Metro), 
                   y = as.numeric(Price)), fill="#FF9999", color="#56B4E9", outlier.size = 0.5) +
  labs(x = "Metro Area", y = "Price", #note these are opposite because I use coord_flip to flip the axes
       colour = "#E0E0E0") +
  coord_flip()

```

## Pricing Distribution for Philadelphia (Box plots)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
price = gather(data = zillow, "Month", "Price", 7:81, factor_key = T)

ggplot(price[price$State == "PA" & !is.na(price$Metro),]) +
  geom_boxplot(aes(x = fct_rev(Metro), 
                   y = as.numeric(Price)), fill="#9999CC", color="#66CC99", outlier.size = 0.5) +
  labs(x = "Metro Area", y = "Price", #note these are opposite because I use coord_flip to flip the axes
       colour = "#E0E0E0") +
  coord_flip()

```

## Dallas county rent prices (point plot across time)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 7}
# price
ggplot(price[price$County == "Dallas" & !is.na(price$City),], 
               aes(y = fct_rev(City), x = Price)) +
  labs(title = "Zillow | Dallas County Rent Prices",
       subtitle = "2010 - 2017", 
       x = "Price", y = "Dallas County", 
       colour = "Time") +
  geom_point(shape = 20, alpha = 0.4, size = 3, aes(color = Month)) +
  scale_color_discrete(l = 45, h = c(30, 330)) 

```

## Los Angeles county rent prices (point plot across time)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 15}

ggplot(price[price$County == "Los Angeles" & !is.na(price$City),], 
               aes(y = fct_rev(City), x = Price)) +
  labs(title = "Zillow | Los Angeles County Rent Prices",
       subtitle = "2010 - 2017", 
       x = "Price", y = "LA County", 
       colour = "Time") +
  geom_point(shape = 16, alpha = 0.4, size = 3, aes(color = Month)) +
  scale_colour_viridis_d()
  # scale_colour_gradientn(colours=rainbow(4))

```

## Average rent pricing with minimum maximum deviation
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 8}
# Create variable of numeric year
price$Year = as.character(price$Month)
price$Year = as.numeric(substr(price$Month, nchar(as.character(price$Month)) - 3, nchar(as.character(price$Month))))

# Calculate range for each state, by year
states = price[!is.na(price$State),] %>%
  group_by(State, Year) %>%
  summarise(Mean = round(mean(Price),0),
            Min = min(Price),
            Max = max(Price)) 

# Plot change over time, by state.
ggplot(states[!is.na(states$Mean),], 
                aes(x = fct_rev(State),
                    y = as.numeric(Mean))) +
  labs(title = "Zillow | Mean US Rent Prices",
       subtitle = "2010 - 2017", 
       x = "States", y = "Average Price", 
       colour = "#E0E0E0") +
  geom_point(shape = 20, alpha = 0.8, size = 5, aes(color = Year)) +
  scale_color_continuous(aes(guide = ""), low = "black", high = "purple") + 
  guides(fill = guide_colourbar(barwidth = 0.7, barheight = 15)) +
  coord_flip() 

```

## Hex plot with number of counties for top 5 cities
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=12}

num_city = 5
values=head(zillow, num_city, )
values=data.frame(t(as.matrix(values[,7:81])))
colnames(values)=zillow[1:num_city,2]

date = seq(as.Date("2010/11/01"), as.Date("2017/01/31"), "month")
date = as.yearmon(date)
ts=zoo(values,order.by = date)
values=fortify(ts)
values$Index=as.Date(values$Index)

autoplot(ts,facets = NULL)+ geom_hex(size=1.5, alpha=0.7) + scale_fill_viridis_c() + 
  guides(fill = guide_colourbar(barwidth = 0.7, barheight = 15)) +
  theme_minimal()+ 
  labs(x="Time",y="Price")
```




