---
title: "6101 Project 1"
author: Jeffrey Hu, Aaron Rock, Sanchit Vijay
date: 9/27/22 to 10/27/22
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# Make sure that install.packages("matrixStats"), install.packages("reshape2"), install.packages("reshape") are installed

library(ezids)
library(ggplot2)
library(tidyverse)
library(MASS)
library(reshape)
library(dplyr)
library(matrixStats)
library(zoo)
library(ggfortify)
library(reshape2)
library(hexbin)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Summary Paper Instructions
Write a roughly 10-page (definitely no more than 4000 words, charts do not count)
summary of the research and EDA process of your project. The summary should be
prepared in Rmarkdown and knitted into HTML. You may take some of the work in part
II (such as graphs and results) to include here. They can overlap. You may also make
modifications to your analysis based on feedback from your presentation. This summary is
to be presented to your boss, your client, or to be submitted for publication in journals.
Potential area of topics to address in this summary may include:
 What do we know about this dataset?
 What are the limitations of the dataset?
 How was the information gathered?
 What analysis has already been completed related to the content in your dataset?
 How did the research you gathered contribute to your question development?
 What additional information would be beneficial?
 How did your question change, if at all, after Exploratory Data Analysis?
 Based on EDA can you begin to sketch out an answer to your question?
 References (APA style preferred)



## Introduction to the dataset
The dataset we explored is on the monthly rental prices of the United States of America from 2010 to 2017 sourced from the application Zillow. This dataset was found on the website Kaggle and has no listed author. The contents of the dataset covers all types of properties such as apartments, townhomes, single family, and multi-family properties that make up the median city code price listing. It is important to note a few important facts about the dataset before going into any more detail. First off, the rental prices are not listing of physical properties but in fact is the median price of the specific city code. Secondly, the dataset does not include every geographical location in the United States and does not include the state of Alaska. Our dataset consists of 13131 observations. There are 1738 rows that are completely empty so after removing these rows we have 11348 observations and 81 variables. Our variables consist of the city code, which is the unique identification of each observation, city, state, metro, country, and population rank. The rest of the variables are time stamps of the monthly rental price of that row’s city code from November 2010 to January 2017. Zillow creates rent forecasting grounded on exclusive statistical and machine learning models also known as Zestimates. Zestimate history only goes back until November 2010 hence why the first time stamp begins on this date. Our dataset was last updated on January 2017 and will be the last variable of the overall dataset. 
  
## Biases and Limitations
Although the dataset has 11348 substantial observations, we need to highlight limitations and biases of the dataset we have chosen to explore. As was stated above the dataset does not include every single zip code in the United States and affected certain results in our EDA. There are a plethora of websites and companies that list rental properties. A major bias is that we have chosen a dataset that limits itself to only using properties that were listed on the Zillow website, so even when talking about the changes of locations and their rental prices it is only considering the information that was listed on Zillow and used for their Zestimate.

## Questions Posed
The goal of exploring the Zillow dataset was to answer and look into three main questions that were introduced about how time and location affected rental prices. The first question is between the east and west coast what region’s price rose the most or were they similar from the first time stamp to the last time stamp? Was there ever overlapping in pricing? Secondly, what cities saw the most change in price differentiation amongst the cities that were picked? Lastly, with the total population …………………….
```{r, echo = FALSE}
#First, load the file
# Import the data
zillow = data.frame(read.csv("price.csv"))
head(zillow)
```
```{r, echo = FALSE}
# Remove blank pricing data
# Remove empty rows, NA's are only present if no data was collected for the location
zillow = na.omit(zillow)
```

```{r, echo = FALSE}
# Subset the Zillow data into east coast and west coast and annual subsets of the regions.
# going through and discarding states that are not defined by the Uited States as West or East coast states

eastcoast = subset(zillow, State == 'CT' | State == 'NY' | State == 'VA' | State == 'FL' | State == 'NJ' | State == 'NC' |State == 'SC' | State == 'GA' | State == 'DE' | State == 'MA' | State == 'MD' | State == 'ME' | State == 'RI' | State == 'NH')
eastcoast_annual = eastcoast[,c(1:81)]
eastcoast_annual = eastcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
eastcoast_annual = eastcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
eastcoast_annual = eastcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
eastcoast_annual = eastcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
eastcoast_annual = eastcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
eastcoast_annual = eastcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
eastcoast_annual = eastcoast_annual[-c(14,15)]

westcoast = subset(zillow, State == 'CA' | State == 'OR' | State == 'WA' | State == 'HI' | State == 'AK')

westcoast_annual = westcoast[,c(1:81)]
westcoast_annual = westcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
westcoast_annual = westcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
westcoast_annual = westcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
westcoast_annual = westcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
westcoast_annual = westcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
westcoast_annual = westcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
westcoast_annual = westcoast_annual[-c(14,15)]

westcoast = westcoast[-c(2,3,4,5,6)]

westAnn = westcoast_annual
westAnn = westAnn[-c(2,3,4,5,6)]

eastAnn = eastcoast_annual
eastAnn = eastAnn[-c(2,3,4,5,6)]
```


```{r, echo = FALSE}
# Find the difference between of rent price and add it as a column AND add average price of rent over the years column
# use dplyr package to create new columns Price.Diff and Avg.Price
eastcoast_annual <- eastcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

westcoast_annual <- westcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

eastcoast_annual <- eastcoast_annual %>%
     mutate(Avg.Price = rowMeans(eastcoast_annual[7:13]))

westcoast_annual <- westcoast_annual %>%
     mutate(Avg.Price = rowMeans(westcoast_annual[7:13]))
```





```{r, echo = FALSE}
# Clean Data of Outliers
# clean the data
westcoast_annualClean = outlierKD2(westcoast_annual, November.2010, rm = TRUE, FALSE,FALSE,FALSE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, November.2016, rm = TRUE, FALSE,FALSE,FALSE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Avg.Price, rm = TRUE, FALSE,FALSE,FALSE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Price.Diff, rm = TRUE, FALSE,FALSE,FALSE)

eastcoast_annualClean = outlierKD2(eastcoast_annual, November.2010, rm = TRUE, FALSE,FALSE,FALSE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, November.2016, rm = TRUE, FALSE,FALSE,FALSE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Avg.Price, rm = TRUE, FALSE,FALSE,FALSE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Price.Diff, rm = TRUE, FALSE,FALSE,FALSE)
```


```{r, echo = FALSE}
# Transpose Data Correctly
rownames(westAnn) = westAnn$City.Code
westAnn = westAnn[c(-1)]
westAnn = as.data.frame(t(westAnn))
names(westAnn)<-str_replace_all(names(westAnn), c(" " = "" , "," = ""))
westAnn <- cbind(Date = rownames(westAnn), westAnn)
rownames(westAnn) <- 1:nrow(westAnn)
head(westAnn)


rownames(eastAnn) = eastAnn$City.Code
eastAnn = eastAnn[c(-1)]
eastAnn = as.data.frame(t(eastAnn))
names(eastAnn)<-str_replace_all(names(eastAnn), c(" " = "" , "," = ""))
eastAnn <- cbind(Date = rownames(eastAnn), eastAnn)
rownames(eastAnn) <- 1:nrow(eastAnn)
head(eastAnn)
```


## East Coast versus West Coast
The first question to be discussed will be the differences and comparisons of the East coast and West coast. Each region is defined by what the United States refers to as East and West coast. The East coast is made up of fourteen states and after subsetting from the original lump data frame has 3988 observations. The West coast is comprised of five states and following the creation of its own data frame the west coast has 1221 observations. We will first look at the frequency of rental prices by viewing the histograms of each region and the month of November with the year 2010 depicted in the light red color and year 2016 shown in the light blue color. For each chart shown there will be a plot including outliers and then a chart not including outliers to help with distribution. 

```{r fig1, fig.height = 5, fig.width = 7, fig.align = "center"}
library(ggplot2)

colors <- c("November.2010" = "red", "November.2016" = "blue")

ggplot(westcoast_annual) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016 with Outliers", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)


ggplot(westcoast_annualClean) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016 without Outliers", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)
```

The most observed rental price for the West coast for both 2010 and 2016 of November sat right around 1250 a month with the bin shifting slightly to the right for the year 2016 with the number of frequencies being above 100 for 2010 and about 75 for the year 2016. Prices for this region did get more expensive and we can notice that for 2016 the plot is crawling up the x-axis showing that prices are climbing.

```{r fig2, fig.height = 5, fig.width = 7, fig.align = "center"}
colors <- c("November.2010" = "red", "November.2016" = "blue")

ggplot(eastcoast_annual) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)

ggplot(eastcoast_annualClean) + 
    geom_histogram(aes(November.2010, fill = "November.2010"),alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016, fill = "November.2016"),alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", 
         y = "Frequency", fill = "Legend") +
    scale_color_manual(values = colors)
```

For the East coast, it is clear again that about 1250 a month for the rental price is the most frequent price. Just like with the West coast we see that the blue, depicting the year 2016, is pulling itself up the x-axis again showing climbing rental prices. One thing to note about the East coast plots is the y-axis clearly being a larger number showing just the differences in observations between the two regions.

When looking at each state in the separate regions when plotted as a boxplot it is clear there is a wide variety of means and interquartile ranges. For the West coast, we see all states for year 2010 and 2016 have overlapping interquartile ranges telling us that for the West coast the prices are somewhat similar for each state. Going back to the limitations of the dataset, it’s obvious when plotted that the west coast holds two states known nationally for having higher rental prices this being Hawaii and California, and although Alaska is defined as a West coast state it was not included in the original dataset. Furthermore, the East coast region when plotted shows interquartile ranges and means that are not always overlapping with the other states in the region. This again highlights some bias of which city codes were chosen for when the original dataset was assembled.
# West Coast 
```{r, figures-side2, fig.show="hold", out.width="50%"}
ggplot(westcoast_annualClean, aes(State, November.2010, color=State)) + geom_boxplot()+labs(title = "West Coast States November 2010 Rent Prices without Outliers")
ggplot(westcoast_annualClean, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "West Coast States November 2016 Rent Prices without Outliers")
```
# East Coast 
```{r, figures-side4, fig.show="hold", out.width="50%"}
ggplot(eastcoast_annualClean, aes(State, November.2010, color=State)) + geom_boxplot()+labs(title = "East Coast States November 2010 Rent Prices without Outliers")
ggplot(eastcoast_annualClean, aes(State, November.2016, color=State)) + geom_boxplot()+labs(title = "East Coast States November 2016 Rent Prices without Outliers")
```

Clearly after plotting the coasts as a whole region in lieu of the coast’s states that the West coast generally has higher rental prices both on the November 2010 plot and on the November 2016 plot. The average of the West coast monthly rental price on November 2010 is 1534 a month with the interquartile ranging from 1171 to 1834 a month. For the November 2016 plot, the average monthly rental price is 1823 a month with a interquartile range spanning from 1294 to 2272 a month. 

```{r, figures-side5, fig.show="hold", out.width="50%"}
ggplot(westcoast_annualClean, aes(x=November.2010)) + geom_boxplot(color="blue", outlier.shape = 9, outlier.size = 2, outlier.color = "blue") + labs(title="Entire West Coast November 2010 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))
ggplot(westcoast_annualClean, aes(x=November.2016)) + geom_boxplot(color="blue", outlier.shape = 9, outlier.size = 2, outlier.color = "blue") + labs(title="Entire West Coast November 2016 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))
```
It is apparent of the price difference between the two regions with the East coast average monthly rental price on November 2010 being 1370 a month with the interquartile range being 1082 to 1598, considerably lower than the 2010 data from the West coast. For the month of November in 2016 in the East coast the average monthly rent is 1498 with a interquartile range comprising of 1170 to 1765, yet again lower than the West's 2016 numbers.

```{r, figures-side6, fig.show="hold", out.width="50%"}
ggplot(eastcoast_annualClean, aes(x=November.2010)) + geom_boxplot(color="red", outlier.shape = 9, outlier.size = 2, outlier.color = "red") + labs(title="Entire East Coast November 2010 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))	
ggplot(eastcoast_annualClean, aes(x=November.2016)) + geom_boxplot(color="red", outlier.shape = 9, outlier.size = 2, outlier.color = "red") + labs(title="Entire East Coast November 2016 Boxplot") + scale_x_continuous(breaks=seq(500, 4000, 500))	
```
Now knowing that the West coast easily coast around 200 to 300 dollars more a month, the last question to answer and to reassure that the West coast region costs more on average is to see if there was ever overlapping in means. This being done by using t-tests on the variables November 2010 and the variable November 2016. The t-test was set at two different intervals one being at 80 and the other being at 99 percent. For the time variable of November 2010 at an 80 percent confidence interval the range spanned from 1515 to 1553 for the West coast and ranged from 1362 to 1378 for the East coast. In 2016, still at 80 percent, the West coast had a range of 1798 to 1849 and the East with a range from 1489 to 1508. When setting the confidence interval at 99 percent the West has a range of 1497 to 1571 for 2010 and 1771 to 1876 for the year 2016. The East coast had a range from 1354 to 1386 for 2010 and a span of 1480 to 1517 for 2016. The results of the t-test endorse the conclusion that the West coast had a higher average monthly rental price and never had overlapping at the 80 or 99 percent confidence interval of the t-tests with their being a differnece again of roughly 200 to 300 dollars a month.


# T Test of First and Last Month of East and  West Coast
```{r, echo = FALSE}
loadPkg("BSDA") # for z.test
west2010_80 = t.test(x=westcoast_annualClean$November.2010, conf.level = 0.80)
west2010_80
west2016_80 = t.test(x=westcoast_annualClean$November.2016, conf.level = 0.80)
west2016_80

west2010_99 = t.test(x=westcoast_annualClean$November.2010, conf.level = 0.99)
west2010_99
west2016_99 = t.test(x=westcoast_annualClean$November.2016, conf.level = 0.99)
west2016_99

east2010_80 = t.test(x=eastcoast_annualClean$November.2010, conf.level = 0.80)
east2010_80
east2016_80 = t.test(x=eastcoast_annualClean$November.2016, conf.level = 0.80)
east2016_80

east2010_99 = t.test(x=eastcoast_annualClean$November.2010, conf.level = 0.99)
east2010_99
east2016_99 = t.test(x=eastcoast_annualClean$November.2016, conf.level = 0.99)
east2016_99
```

\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
Start of Cities of Interest File
\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\


## Load the file
```{r}
zillow_prices = data.frame(read.csv("price.csv"))
zillow_focus = zillow_prices[c(1:5,7,13,17,23,24,26,113,162),]
zillow_focus = zillow_focus[c(-1,-3,-4,-5,-6)]
zillow_focus
```


## Clean and Transpose
Our cities of interest are NYC (Queens) (only data from December 2011 Onwards), LA, PA, Houston (Harris), Chicago (Cook), Dallas, Las Vegas, San Fran, Detroit (Wayne), Roanoke, Richmond, District of Columbia, Baltimore, Seattle(King). Here, we clean data set and transpose the dates/cities
```{r, results="markup"}
rownames(zillow_focus) = zillow_focus$City
zillow_focus = zillow_focus[c(-1)]
zillow_focus = as.data.frame(t(zillow_focus))
names(zillow_focus)<-str_replace_all(names(zillow_focus), c(" " = "" , "," = ""))
head(zillow_focus)
str(zillow_focus)
```

Turn date into a column
```{r, results="markup"}
zillow_focus <- cbind(Date = rownames(zillow_focus), zillow_focus)
rownames(zillow_focus) <- 1:nrow(zillow_focus)
head(zillow_focus)
```


## Convert the Date column from chr to date, and Total
Here, we are converting strings to dates and to %d/%m/%Y form with 01 for %d
```{r, results="markup"}
zillow_focus_test = zillow_focus
zillow_focus_test$Date <- str_replace_all(zillow_focus_test$Date, "[.]", " ")
zillow_focus_test$Month <- str_extract(zillow_focus_test$Date, "(\\w+)")
zillow_focus_test$Year <- str_extract(zillow_focus_test$Date, "\\w+$")
tmp_date = paste(match(zillow_focus_test$Month, month.name), zillow_focus_test$Year, sep="/")
tmp_date = paste("01", tmp_date, sep="/")
tmp_date = as.Date(tmp_date, "%d/%m/%Y")
zillow_focus$Date <- tmp_date
head(zillow_focus)
str(zillow_focus)
```


Add a total column, not including New York (which has NAs)
```{r, results="markup"}
zillow_total = zillow_focus
zillow_total$Total <- rowSums(zillow_total[3:14])
head(zillow_total)
```


## Line plot of total
Line plots of time series, one with only time
```{r}
ggplot(zillow_total, aes(x=Date, y=Total)) + geom_point() + geom_line() + labs(x="Time", y="Total Rent Prices", title="Plot of City Rent Change over Time")
```


## Plots of the Cities and Prices
Melt the ggplot by the cities
```{r, results="markup"}
zillow_melt <- melt(zillow_focus, id = "Date")
names(zillow_melt)[2] <- "City"
names(zillow_melt)[3] <- "Prices"
head(zillow_melt)
```


I'm not sure if histograms would be meaningful in the context of my part, but here's one of all the cities melted together, ignoring New York again. We can do more if you guys believe it's necessary. There is a huge skew because the bigger/more expensive cities chosen would have higher rent prices due to a difference in the cost of living.
I'll also do the Mean/Median/Quartiles of the overall plots here.
```{r, results="markup"}
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_histogram(binwidth=300)
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_boxplot()
summary(zillow_melt$Prices)
```

Lines of the time series over their changes.
```{r}
ggplot(zillow_melt, aes(x=Date, y=Prices, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Rent Price", title="Plot of City Rent Prices over Time")
```


## First Order change plots
Find the first order change, difference between each month
```{r, results="markup"}
zillow_change <- as.data.frame(lapply(zillow_focus[2:14], diff, lag=1))
zillow_change$Date <- zillow_focus$Date[2:75]
zillow_change <- zillow_change %>% dplyr::select(Date, everything())
head(zillow_change)
```


Melt, and line graph of the changes for each city.
The graph is kind of hard to visualize, but it looks like San Francisco had the highest change from time to time.
```{r}
zillow_cmelt <- melt(zillow_change, id = "Date")
names(zillow_cmelt)[2] <- "City"
names(zillow_cmelt)[3] <- "PriceChange"
ggplot(zillow_cmelt, aes(x=Date, y=PriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Price Change", title="Plot of City Rent Change over Time")
```

Cumulative change, to make the trends more clear in respect to each other
```{r, results="markup"}
zillow_cumulative <- as.data.frame(cumsum(zillow_change[2:14]))
zillow_cumulative$Date <- zillow_focus$Date[2:75]
zillow_cumulative <- zillow_cumulative %>% dplyr::select(Date, everything())
head(zillow_cumulative)
```


Another graph, this time of cumulative change. San Francisco had the greatest change, while Seattle had the second largest. Detroit had the most negative change, falling below 0 overall.
```{r}
zillow_cumumelt <- melt(zillow_cumulative, id = "Date")
names(zillow_cumumelt)[2] <- "City"
names(zillow_cumumelt)[3] <- "CumulativePriceChange"
ggplot(zillow_cumumelt, aes(x=Date, y=CumulativePriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumulative Change", title="Plot of Cumulative City Rent Change over Time")
```


## First Order change, as a percentage proportion of the total rent
Find the first order change as a Percent
```{r, results="markup"}
zillow_proportion <- zillow_change[2:14]/zillow_focus[c(2:75), 2:14]*100
zillow_proportion$Date <- zillow_focus$Date[2:75]
zillow_proportion <- zillow_proportion %>% dplyr::select(Date, everything())
head(zillow_proportion)
```

It looks like Richmond has a pretty varied proportional change. It's still hard to make anything out of it, so let's move to the cumulative proportional change again.
```{r}
zillow_pmelt <- melt(zillow_proportion, id = "Date")
names(zillow_pmelt)[2] <- "City"
names(zillow_pmelt)[3] <- "ProportionChange"
ggplot(zillow_pmelt, aes(x=Date, y=ProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Proportional Change", title="Plot of City Rent Change over Time")
```


Quick histogram of the distribution of proportional changes. The proportional changes over time look pretty normal taken together as a whole, with maybe a slight right skew, which makes sense as we would guess that rent prices should generally increase over time with inflation and what not. Remember, these changes are in percentages. There are outliers on both sides of the boxplot, which matches what we see in the histogram. Here are the measures of variance and qqnorm plots as well
```{r, results="markup"}
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_histogram() + labs(x = "Proportional Change in Percentages")
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_boxplot()
qqnorm(zillow_pmelt$ProportionChange)
summary(zillow_pmelt$ProportionChange)
var(zillow_pmelt$ProportionChange, na.rm=TRUE)
sd(zillow_pmelt$ProportionChange, na.rm=TRUE)
```

Just making the cumulative melt for proportions now. We will drop New York before plotting the values, as there are too many NAs.

```{r, results="markup"}
zillow_pdrop <- zillow_proportion[3:14]
zillow_pcumulative <- as.data.frame(cumsum(zillow_pdrop))
zillow_pcumulative$Date <- zillow_focus$Date[2:75]
zillow_pcumulative <- zillow_pcumulative %>% dplyr::select(Date, everything())
head(zillow_pcumulative)
```

We will drop New York before plotting the values, as there are too many NAs.
Now, it looks like the proportional change of Seattle is a bit higher than San Francisco, and Richmond is up there too. Detroit still has the most negative proportional change overall.

```{r}
zillow_pcumumelt <- melt(zillow_pcumulative, id = "Date")
names(zillow_pcumumelt)[2] <- "City"
names(zillow_pcumumelt)[3] <- "CumulativeProportionChange"
ggplot(zillow_pcumumelt, aes(x=Date, y=CumulativeProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumu. Prop. Change", title="Plot of Cumulative City Rent Change over Time")
```

Box plot of each city's proportional change. As expected, San Francisco and Roanoke seem to have the largest ranges and variances. We are also dropping New York again.
```{r, fig.align="center", fig.width = 10, fig.height=5}
zillow_prodrop = zillow_proportion[-2]
zillow_pdmelt <- melt(zillow_prodrop, id = "Date")
names(zillow_pdmelt)[2] <- "City"
names(zillow_pdmelt)[3] <- "ProportionChange"
ggplot(zillow_pdmelt, aes(x=City, y=ProportionChange, color=City, na.rm=TRUE)) + geom_boxplot(outlier.shape=2, outlier.size=1)
```


## Anova Test and Post-hoc Tukey HSD
We will now anova test the proportional changes against each other, and see if the average change between years is different. Since we have a p-value less than our alpha, there are significant differences in changes, and thus, we look at the Post-hoc Tukey HSD. It looks like the pairs that have significantly different average proportional changes are [Detroit-LosAngeles], [Roanoke-LosAngeles], [SanFrancisco-Chicago], [Seattle-Chicago], [Richmond-Chicago], [Detroit-Houston], [Seattle-Philadelphia], [SanFrancisco-LasVegas], [Seattle-LasVegas], [Richmond-LasVegas], [Detroit-SanFrancisco], [Roanoke-SanFrancisco], [Seattle-Detroit], [Washington-Detroit], [Richmond-Detroit], [Baltimore-Seattle], [Roanoke-Seattle], and finally, [Roanoke-Richmond].
```{r, results="markup"}
pm_anova = aov(zillow_pdmelt$ProportionChange ~ zillow_pdmelt$City)
summary(pm_anova)
tukeyAoV <- TukeyHSD(pm_anova)
tukeyAoV
```

\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
Start of Total Population File
\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\

## State wise countplot  
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
ggplot(zillow, aes(x=State)) + geom_bar(colour="blue", fill="purple", alpha=0.6)
```

## Time-series plot for top 10 cities  
```{r, results='markup', fig.align="center", fig.width = 12, fig.height=5}
num_city = 10
values=head(zillow, num_city, )
values=data.frame(t(as.matrix(values[,7:81])))
colnames(values)=zillow[1:num_city,2]

date = seq(as.Date("2010/11/01"), as.Date("2017/01/31"), "month")
date = as.yearmon(date)
ts=zoo(values,order.by = date)
values=fortify(ts)
values$Index=as.Date(values$Index)

autoplot(ts,facets = NULL)+ geom_point(size=0.5) +
  theme_minimal()+
  labs(x="Time",y="Price")

```

## Pricing Distribution for California (Box plots)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
price = gather(data = zillow, "Month", "Price", 7:81, factor_key = T)

ggplot(price[price$State == "CA" & !is.na(price$Metro),]) +
  geom_boxplot(aes(x = fct_rev(Metro), 
                   y = as.numeric(Price)), fill="#FF9999", color="#56B4E9", outlier.size = 0.5) +
  labs(x = "Metro Area", y = "Price", #note these are opposite because I use coord_flip to flip the axes
       colour = "#E0E0E0") +
  coord_flip()

```

## Pricing Distribution for Philadelphia (Box plots)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=5}
price = gather(data = zillow, "Month", "Price", 7:81, factor_key = T)

ggplot(price[price$State == "PA" & !is.na(price$Metro),]) +
  geom_boxplot(aes(x = fct_rev(Metro), 
                   y = as.numeric(Price)), fill="#9999CC", color="#66CC99", outlier.size = 0.5) +
  labs(x = "Metro Area", y = "Price", #note these are opposite because I use coord_flip to flip the axes
       colour = "#E0E0E0") +
  coord_flip()

```

## Dallas county rent prices (point plot across time)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 7}
# price
ggplot(price[price$County == "Dallas" & !is.na(price$City),], 
               aes(y = fct_rev(City), x = Price)) +
  labs(title = "Zillow | Dallas County Rent Prices",
       subtitle = "2010 - 2017", 
       x = "Price", y = "Dallas County", 
       colour = "Time") +
  geom_point(shape = 20, alpha = 0.4, size = 3, aes(color = Month)) +
  scale_color_discrete(l = 45, h = c(30, 330)) 

```

## Los Angeles county rent prices (point plot across time)
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 15}

ggplot(price[price$County == "Los Angeles" & !is.na(price$City),], 
               aes(y = fct_rev(City), x = Price)) +
  labs(title = "Zillow | Los Angeles County Rent Prices",
       subtitle = "2010 - 2017", 
       x = "Price", y = "LA County", 
       colour = "Time") +
  geom_point(shape = 16, alpha = 0.4, size = 3, aes(color = Month)) +
  scale_colour_viridis_d()
  # scale_colour_gradientn(colours=rainbow(4))

```

## Average rent pricing with minimum maximum deviation
```{r, results='markup', fig.align="center", fig.width = 10, fig.height = 8}
# Create variable of numeric year
price$Year = as.character(price$Month)
price$Year = as.numeric(substr(price$Month, nchar(as.character(price$Month)) - 3, nchar(as.character(price$Month))))

# Calculate range for each state, by year
states = price[!is.na(price$State),] %>%
  group_by(State, Year) %>%
  summarise(Mean = round(mean(Price),0),
            Min = min(Price),
            Max = max(Price)) 

# Plot change over time, by state.
ggplot(states[!is.na(states$Mean),], 
                aes(x = fct_rev(State),
                    y = as.numeric(Mean))) +
  labs(title = "Zillow | Mean US Rent Prices",
       subtitle = "2010 - 2017", 
       x = "States", y = "Average Price", 
       colour = "#E0E0E0") +
  geom_point(shape = 20, alpha = 0.8, size = 5, aes(color = Year)) +
  scale_color_continuous(aes(guide = ""), low = "black", high = "purple") + 
  guides(fill = guide_colourbar(barwidth = 0.7, barheight = 15)) +
  coord_flip() 

```

## Hex plot with number of counties for top 5 cities
```{r, results='markup', fig.align="center", fig.width = 10, fig.height=12}

num_city = 5
values=head(zillow, num_city, )
values=data.frame(t(as.matrix(values[,7:81])))
colnames(values)=zillow[1:num_city,2]

date = seq(as.Date("2010/11/01"), as.Date("2017/01/31"), "month")
date = as.yearmon(date)
ts=zoo(values,order.by = date)
values=fortify(ts)
values$Index=as.Date(values$Index)

autoplot(ts,facets = NULL)+ geom_hex(size=1.5, alpha=0.7) + scale_fill_viridis_c() + 
  guides(fill = guide_colourbar(barwidth = 0.7, barheight = 15)) +
  theme_minimal()+ 
  labs(x="Time",y="Price")
```




