---
title: "EDA-Aaron Project 1"
author: "Aaron Rock"
date: "2022-10-19"
output: html_document
---
```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(tidyverse)
library(ggplot2)
library(dplyr)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


```{r}
#First, load the file
zillow = data.frame(read.csv("price.csv"))
head(zillow)
```

Remove blank pricing data
``` {r}
# Remove empty rows
zillow = na.omit(zillow)
```

Subset the Zillow data into east coast and west coast and annual subsets of the regions.
``` {r}
eastcoast = subset(zillow, State == 'CT' | State == 'NY' | State == 'VA' | State == 'FL' | State == 'NJ' | State == 'NC' |State == 'SC' | State == 'GA' | State == 'DE' | State == 'MA' | State == 'MD' | State == 'ME' | State == 'RI' | State == 'NH')
eastcoast_annual = eastcoast[,c(1:81)]
eastcoast_annual = eastcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
eastcoast_annual = eastcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
eastcoast_annual = eastcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
eastcoast_annual = eastcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
eastcoast_annual = eastcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
eastcoast_annual = eastcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
eastcoast_annual = eastcoast_annual[-c(14,15)]

westcoast = subset(zillow, State == 'CA' | State == 'OR' | State == 'WA' | State == 'HI' | State == 'AK')

westcoast_annual = westcoast[,c(1:81)]
westcoast_annual = westcoast_annual[-c(8,9,10,11,12,13,14,15,16,17,18)]
westcoast_annual = westcoast_annual[-c(9,10,11,12,13,14,15,16,17,18,19)]
westcoast_annual = westcoast_annual[-c(10,11,12,13,14,15,16,17,18,19,20)]
westcoast_annual = westcoast_annual[-c(11,12,13,14,15,16,17,18,19,20,21)]
westcoast_annual = westcoast_annual[-c(12,13,14,15,16,17,18,19,20,21,22)]
westcoast_annual = westcoast_annual[-c(13,14,15,16,17,18,19,20,21,22,23)]
westcoast_annual = westcoast_annual[-c(14,15)]

westcoast = westcoast[-c(2,3,4,5,6)]


## Find the difference between of rent price and add it as a column AND add average price of rent over the years column 
eastcoast_annual <- eastcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

westcoast_annual <- westcoast_annual %>%
    mutate(Price.Diff = November.2016 - November.2010)

eastcoast_annual <- eastcoast_annual %>%
     mutate(Avg.Price = rowMeans(eastcoast_annual[7:13]))

westcoast_annual <- westcoast_annual %>%
     mutate(Avg.Price = rowMeans(westcoast_annual[7:13]))
```




## Clean Data of Outliers
```{r,results='asis'}
westcoast_annualClean = outlierKD2(westcoast_annual, November.2010, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, November.2016, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Avg.Price, rm = TRUE)
westcoast_annualClean = outlierKD2(westcoast_annualClean, Price.Diff, rm = TRUE)

eastcoast_annualClean = outlierKD2(eastcoast_annual, November.2010, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, November.2016, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Avg.Price, rm = TRUE)
eastcoast_annualClean = outlierKD2(eastcoast_annualClean, Price.Diff, rm = TRUE)
```

## Transpose Data Correctly
```{r, results='asis'}
rownames(westcoast) = westcoast$City.Code
westcoast = westcoast[c(-1)]
westcoast = as.data.frame(t(westcoast))
names(westcoast)<-str_replace_all(names(westcoast), c(" " = "" , "," = ""))
head(westcoast)
westcoast <- cbind(Date = rownames(westcoast), westcoast)
rownames(westcoast) <- 1:nrow(westcoast)
head(westcoast)
```



Summary of the two regions annually
``` {r, echo=T}
summary(eastcoast_annual[7:15])
summary(westcoast_annual[7:15])
xkablesummary(westcoast_annual[7:15], title="West Coast Summary Annually")
xkablesummary(eastcoast_annual[7:15], title="East Coast Summary Annually")
```
## Finding the Outliers to Help with Graphs
```{r, results='asis'}
outWest2010 <- boxplot.stats(westcoast_annual$November.2010)$out
outWest2016 <- boxplot.stats(westcoast_annual$November.2016)$out
outEast2010 <- boxplot.stats(eastcoast_annual$November.2010)$out
outEast2016 <- boxplot.stats(eastcoast_annual$November.2016)$out
min(outWest2010)
min(outWest2016)
min(outEast2010)
min(outEast2016)
```

## West Coast Annual Novemeber 2010 Start and November 2016 End Histogram
```{r}
library(ggplot2)
ggplot(westcoast_annualClean, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2010", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2010", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2016", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annualClean, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 70) + 
    labs(title = "Rental Price Count on the West Coast in November 2016", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annual) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 70) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 70) +
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")

ggplot(westcoast_annualClean) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 40) + 
    labs(title = "Rental Price Count on the West Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")
```

## East Coast Annual Novemeber 2010 Start and November 2016 End Histogram
```{r}
ggplot(eastcoast_annual, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annualClean, aes(November.2010)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annual, aes(November.2016)) + geom_histogram(fill="red",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2010", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annualClean, aes(November.2016)) + geom_histogram(fill="blue",alpha = 0.4,bins= 40) +  
    labs(title = "Rental Price Count on the East Coast from November 2016", x = "Rental Prices", y = "Frequency")

ggplot(eastcoast_annual) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 40) +
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")


ggplot(eastcoast_annualClean) + 
    geom_histogram(aes(November.2010), fill = "red", alpha = 0.4, bins = 40) + 
    geom_histogram(aes(November.2016), fill = "blue", alpha = 0.4, bins = 40) +
    labs(title = "Rental Price Count on the East Coast from November 2010 to November 2016", x = "Rental Prices", y = "Frequency")


```

## Boxplot of West Coast Zillow Rental Prices 2010-2016
```{r}
ggplot(westcoast_annual, aes(State, November.2010)) + geom_boxplot()
ggplot(westcoast_annualClean, aes(State, November.2010)) + geom_boxplot()
ggplot(westcoast_annualClean, aes(x=November.2010)) + geom_boxplot(outlier.shape = 9, outlier.size = 2, outlier.color = "blue")
ggplot(westcoast_annual, aes(State, November.2016)) + geom_boxplot()
ggplot(westcoast_annualClean, aes(State, November.2016)) + geom_boxplot()
ggplot(westcoast_annualClean, aes(x=November.2016)) + geom_boxplot(outlier.shape = 9, outlier.size = 2, outlier.color = "blue")
```

## Boxplot of East Coast Zillow Rental Prices 2010-2016
```{r}
ggplot(eastcoast_annual, aes(State, November.2010)) + geom_boxplot()
ggplot(eastcoast_annualClean, aes(State, November.2010)) + geom_boxplot()
ggplot(eastcoast_annualClean, aes(x=November.2010)) + geom_boxplot(outlier.shape = 9, outlier.size = 2, outlier.color = "red")
ggplot(eastcoast_annual, aes(State, November.2016)) + geom_boxplot()
ggplot(eastcoast_annualClean, aes(State, November.2016)) + geom_boxplot()
ggplot(eastcoast_annualClean, aes(x=November.2016)) + geom_boxplot(outlier.shape = 9, outlier.size = 2, outlier.color = "red")
```

## Scatter plot of West Coast
```{r, results='asis'}
ggplot(westcoast_annual, aes(November.2010,November.2016, color=State)) + geom_point()
ggplot(westcoast_annualClean, aes(November.2010,November.2016, color=State)) + geom_point()
```

## Scatter plot of East Coast
```{r, results='asis'}
ggplot(eastcoast_annual, aes(November.2010,November.2016, color=State)) + geom_point()
ggplot(eastcoast_annualClean, aes(November.2010,November.2016, color=State)) + geom_point()
```

## QQ Plots
```{r, results='asis'}
qqnorm(westcoast_annualClean$November.2010)
qqline(westcoast_annualClean$November.2010)

qqnorm(westcoast_annualClean$November.2016)
qqline(westcoast_annualClean$November.2016)

qqnorm(eastcoast_annualClean$November.2010)
qqline(eastcoast_annualClean$November.2010)

qqnorm(eastcoast_annualClean$November.2016)
qqline(eastcoast_annualClean$November.2016)
```